RUPEE: A fast and accurate purely geometric protein structure search

### 2019-08-22 docs/rupee.tex
intro

In \cite{Ayoub2019} we introduced RUPEE, a purely geometric protein structure search with no dependence on sequences, clustering, and pre-calculated results. 
We compared our results with the mTM-align structure search \cite{Dong2018}, the secondary structure matching (SSM) search \cite{Krissinel2004}, and the CATHEDRAL structural scan \cite{Redfern2007}, and found RUPEE is equal to or better than the best available structure searches on a benchmark of known protein domains. 
Additionally, we showed RUPEE, on average, is able to return results faster than mTM-align and CATHEDRAL.

Since the release of RUPEE, we have observed that the primary usage pattern for RUPEE has been to upload protein structures that have a very low maximum similarity to other structures in the PDB.
In most cases, these uploaded protein structures are the output of a protein structure prediction protocol. 
While in \cite{Ayoub2019} we recognized the desirability of a purely geometric protein structure search, we did not anticipate the utility of such a search to aid in the development of protein structure prediction methods.
For instance, besides measuring the similarity of a predicted structure to a target structure in order to evaluate a protein structure prediction method in development, it is also natural to determine what known structure a predicted structure resembles most.

[got here in my second pass]
Recognizing the need for being able to search on query structures with low maximum similarity to known structures, we have added an additional search mode to RUPEE with increased sensitivity. 
While top-aligned mode, introduced in \cite{Ayoub2019}, is comparitively good for low similarity searches, the all-aligned mode we introduce below provides a significant improvement over top-aligned mode in that case.  
In this paper, we again compare RUPEE's results mTM-align, SSM, and CATHEDRAL but this time we do so for all-aligned mode and we derive a benchmark from free-modelling predictions in CASP12. 
Whereas in \cite{Ayoub2019} we showed that RUPEE is equal or better than those we compared to for a benchmark of known protein structures, in this paper we show that RUPEE, in all-aligned mode, is clearly better than those we compare to using a benchmark drawn from the structures output by protein structure prediction protocols.

@@

Together, all-aligned mode and containment searches are siginificant additions to RUPEE that increase its utility and further demonstrates the advantage of a purely geomietric protein structure search without dependencies on sequence, precalculated results or clustering.

@@

We first provide a brief outline of our linear encoding of protein structures described in more detail in \cite{Ayoub2019}, which still remains at the core of the RUPEE protein structure search. 
Then, we describe the types of similarity scores used for the initial filtering in the different search modes, the addition of search types, changes to top-aligned mode, and the addition of all-aligned mode, in that order.

>>>>

In \cite{Ayoub2019} we introduced RUPEE, a purely geometric protein structure search with no dependence on sequences, clustering, or pre-calculated results. 
We compared our results with the mTM-align structure search \cite{Dong2018}, the secondary structure matching (SSM) search \cite{Krissinel2004}, and the CATHEDRAL structural scan \cite{Redfern2007}, and found RUPEE is equal to or better than the best available structure searches on a benchmark of known protein domains. 
Additionally, we showed RUPEE, on average, is able to return results faster than mTM-align and CATHEDRAL.

Since the release of RUPEE, we have observed that the primary usage for RUPEE has been for uploaded protein structures that have a low maximum similarity to other structures in the PDB.
In most cases, these uploaded protein structures are the output of a protein structure prediction method. 
While in \cite{Ayoub2019} we recognized the desirability of a purely geometric protein structure search, we did not anticipate the utility of such a search to aid in the development of protein structure prediction methods.
For instance, besides measuring the similarity of a predicted structure to a target structure in order to evaluate a protein structure prediction method in development, it is also natural to determine what known structures a predicted structure resembles most.

Recognizing the need to search on structures with low maximum similarity to known structures, we have added an additional search mode to RUPEE with increased sensitivity. 
While top-aligned mode, introduced in \cite{Ayoub2019}, is also good for low similarity searches, the all-aligned mode we introduce below is a significant improvement over top-aligned mode in that case.

In this paper, we again compare the results of RUPEE against mTM-align, SSM, and CATHEDRAL, but this time we do so for all-aligned mode.
Furthermore, we derive a new benchmark for comparisons from protein structure predictions for free-modelling targets in CASP12 available at the CASP web site. 
Whereas in \cite{Ayoub2019} we showed that RUPEE is equal or better than those we compared to for a benchmark of known protein structures, here we show that RUPEE, in all-aligned mode, is clearly better than those we compare to using a benchmark drawn from the structures output by protein structure prediction methods.

@@

Together, all-aligned mode and containment searches are significant additions to RUPEE that increase its utility and further demonstrates the advantage of a purely geometric protein structure search with no dependence on sequences, clustering, or pre-calculated results.

@@

We first give a brief outline of our linear encoding of protein structures described in more detail in \cite{Ayoub2019}, which still remains at the core of the RUPEE protein structure search. 
Then, we describe the types of similarity scores used for the initial filtering in the different search modes, the addition of search types, changes to top-aligned mode, and the addition of all-aligned mode, in that order.
